### 2024-09-26 outputs/paper/paper.tex
made changes on text format, added bullet points

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Analyzing Toronto Island Ferry Ticket Counts},
  pdfauthor={Yingke He},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Analyzing Toronto Island Ferry Ticket Counts\thanks{Code and data
are available at: https://github.com/ohyykk/ferrySale.git}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{A Statistical Exploration of Sales Trends}
\author{Yingke He}
\date{September 23, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study examines the ticket redemption and sales counts for the
Toronto Island Ferry. We analyze time-based trends, relationships
between sales and redemptions, and develop statistical models to predict
ferry ticket usage. The results provide insights into usage patterns,
highlight operational considerations, and suggest avenues for future
research. Our analysis reveals distinct seasonal peaks and troughs in
ferry usage, correlating with tourism and weather conditions.
Furthermore, the study identifies discrepancies between ticket sales and
redemptions, suggesting potential improvements in ticketing strategies
and resource allocation.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}

\section{Introduction}\label{introduction}

The Toronto Island Ferry system has long been an essential
transportation link between downtown Toronto and the Toronto Islands,
serving both residents and tourists. With roots tracing back to the
early days of the city's development, the ferry system reflects both
historical and contemporary mobility patterns in the region. Historical
accounts provided by Varga (1987) highlight the evolution of the Toronto
Islands' natural environment and its significance to the area's
biodiversity and recreation. Additionally, the potential role of ferries
in urban transit systems, as explored by Kopystynski (1979), underscores
their importance in enhancing urban mobility and connectivity.

This paper aims to analyze the ticket counts for the Toronto Island
Ferry, exploring trends in ticket redemption and sales, and drawing
connections between historical developments and current usage patterns.
By integrating insights from both historical and modern perspectives,
including contemporary studies on ferry commuting from Roseman (2020),
this analysis seeks to provide a comprehensive understanding of the
factors influencing ferry usage. The remainder of this paper is
structured as follows: Section 2 describes the dataset used for
analysis, Section 3 outlines the modeling approach, Section 4 presents
the results, and Section 5 discusses the implications of the findings
and potential directions for future research. We use R Core Team (2023)
and Wickham et al. (2019a), and We use Section~\ref{sec-data} to
reference the data section.

\section{Data}\label{sec-data}

\subsection{Overview}\label{sec-data-overview}

The dataset used in this analysis is the 2024 installment of ``Toronto
Island Ferry Ticket Counts'' obtained from OpenDataToronto (Varga
(1987)). This dataset contains information about the number of ferry
tickets redeemed and sold for the Toronto Island Ferry system, which
serves as a critical transportation link between downtown Toronto and
the Toronto Islands.

The dataset is updated regularly and includes time-stamped entries for
ticket redemptions and sales, allowing for a detailed analysis of ferry
usage patterns throughout the year. The data is considered ``open data''
and can be utilized for various analytical purposes, provided that
appropriate attribution is given..

The primary variables included in this analysis are:

\textbf{Timestamp}: The date and time when each ticket count was
recorded.\\
\textbf{Redemption.Count}: The number of ferry tickets redeemed during
the specified time period.\\
\textbf{Sales.Count}: The number of ferry tickets sold during the
specified time period.

The data was processed and analyzed using the R programming language (R
Core Team (2023)). The tidyverse (Wickham et al. (2019b)) and
opendatatoronto packages were employed to download, clean, and prepare
the dataset for analysis. The resulting cleaned dataset was then
utilized to generate the visualizations and statistical analyses
presented in this paper using a collection of libraries which included
the following packages:

\begin{itemize}
\tightlist
\item
  ggplot2 (Wickham 2016)
\item
  knitr (Xie 2023)
\item
  readr (Wickham, Hester, and Bryan 2023)
\end{itemize}

\subsection{Results}\label{sec-data-resultsk}

The table Table~\ref{tbl-cleaned_data} provides a detailed snapshot of
the Toronto Island Ferry ticket transactions for a specific time period.
It includes a unique ID for each transaction, the Time at which each
transaction occurred, the number of Tickets Redeemed, the number of
Tickets Sold, and the Date of the transaction.

\begin{longtable}[]{@{}ccccc@{}}

\caption{\label{tbl-cleaned_data}Sample of cleaned Ferry Ticket Counts
Data}

\tabularnewline

\toprule\noalign{}
ID & Time & Tickets Redeemed & Tickets Sold & Date \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
51 & 23:30:00 & 5 & 4 & 2024-09-24 \\
52 & 23:15:00 & 1 & 3 & 2024-09-24 \\
53 & 23:00:00 & 2 & 0 & 2024-09-24 \\
54 & 22:45:00 & 7 & 3 & 2024-09-24 \\
55 & 22:30:00 & 11 & 0 & 2024-09-24 \\
56 & 22:15:00 & 2 & 1 & 2024-09-24 \\
57 & 22:00:00 & 5 & 6 & 2024-09-24 \\
58 & 21:45:00 & 2 & 3 & 2024-09-24 \\
59 & 21:30:00 & 14 & 41 & 2024-09-24 \\
60 & 21:15:00 & 3 & 1 & 2024-09-24 \\

\end{longtable}

Sample Cleaned Data

This table shows the typical flucuations in ticket sales and redemptions
throughout the day. For example, during late evening hours, we observe
varying ticket redemption and sales patterns, which may correspond to
changes in passenger demand as the ferry service approaches the end of
its operational hours.

\subsection{Redemption and Sales Counts Over
Time}\label{redemption-and-sales-counts-over-time}

To gain an initial insight into the dataset, we plotted the trends in
ticket redemption and sales counts over time. The line plot shown in
Figure~\ref{fig-linePlot} provides a clear visual representation of
these trends, serving as a starting point for deeper analysis of ferry
usage patterns.

The dataset used in this analysis includes timestamped records of ticket
redemptions and sales for the Toronto Island Ferry system, offering a
time series perspective on ferry usage. The plot in
Figure~\ref{fig-linePlot} specifically highlights how these counts
fluctuate over time, reflecting the flow of passenger demand. The y-axis
for redemption counts is displayed in blue, while the y-axis for sales
counts is displayed in green, but inverted to better show their
correlation.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-linePlot-1.pdf}

}

\caption{\label{fig-linePlot}Redemption and Sales Counts Over Time}

\end{figure}%

As shown in Figure~\ref{fig-linePlot}, the x-axis denotes the timestamps
of each record, while the y-axis represents the number of tickets
processed. The blue line corresponds to the redemption counts, and the
green line represents the sales counts, plotted inversely to emphasize
the contrast between the two trends. The symmetrical y-axis allows for a
clearer comparison between peaks and troughs in both counts over time.

The plot reveals significant variations in ferry usage over time. We
observe several peaks in both redemption and sales counts, likely
aligning with periods of high activity, such as weekends, holidays, or
the peak tourist season during the summer months. These spikes suggest
increased demand during favorable weather or event-driven occasions.
Conversely, the noticeable dips in the graph may correspond to regular
weekdays or off-season times, indicating lower ferry usage during these
periods. This information is valuable for understanding the temporal
dynamics of ferry operations and can inform better resource management
and scheduling.

\subsection{Redemption vs Sales Count}\label{redemption-vs-sales-count}

The scatter plot in Figure~\ref{fig-scatterPlot} is designed to explore
the relationship between ticket redemption and sales counts for the
Toronto Island Ferry system. This visualization allows us to understand
whether there is a correlation between the number of tickets sold and
the number of tickets redeemed during the same time period.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-scatterPlot-1.pdf}

}

\caption{\label{fig-scatterPlot}Scatter Plot of Redemption vs Sales
Count}

\end{figure}%

Figure~\ref{fig-scatterPlot} displays each data point as a purple dot,
where the y-axis represents the count of tickets redeemed, and the
x-axis denotes the count of tickets sold. This plot provides insight
into the degree of alignment between ticket sales and actual usage
(redemptions).

The scatter plot shows how ticket sales and redemptions relate to each
other over time. If most points lie along a diagonal line, it indicates
a strong positive correlation between the two variables, suggesting that
a high number of tickets sold typically corresponds to a high number of
tickets redeemed. This scenario could imply effective ticket
utilization, where most purchased tickets are being used for travel.

\newpage

\subsection{Side by Side Comparison}\label{side-by-side-comparison}

The box plot shown in Figure~\ref{fig-boxPlot} provides a summary of the
distribution of ticket sales and redemption counts for the Toronto
Island Ferry system. This visualization helps to identify the spread,
central tendency, and any potential outliers in the data for both
categories, offering a comparative view of the sales and redemption
behaviors.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-boxPlot-1.pdf}

}

\caption{\label{fig-boxPlot}Box Plot of Sales and Redemption Counts}

\end{figure}%

Figure~\ref{fig-boxPlot} illustrates two box plots side by side: one for
sales counts (light blue) and the other for redemption counts (light
green). Each box plot provides a visual summary of the data
distribution, highlighting the median, quartiles, and any outliers for
each category.

The box plots reveal key differences and similarities in the
distribution of sales and redemption counts. The horizontal line inside
each box represents the median count, providing a measure of central
tendency. The top and bottom edges of the box denote the interquartile
range (IQR), which captures the middle 50\% of the data. The
``whiskers'' extending from the box show the range of values within 1.5
times the IQR from the first and third quartiles, while any points
beyond these whiskers are considered outliers.

\section{Discussion}\label{sec-discussion}

In Section 2.2, the time series analysis of ticket sales and redemptions
(Figure Figure~\ref{fig-linePlot}) revealed clear seasonal patterns,
with pronounced peaks during the warmer months. This trend aligns with
the historical and cultural significance of the Toronto Islands as a
recreational destination, as documented by Varga (1987). The increased
demand during these periods suggests that the ferry system plays a
crucial role in supporting both tourism and local leisure activities.
These seasonal fluctuations underscore the need for effective resource
management and scheduling to accommodate surges in passenger volume
during peak times.

The scatter plot analysis in Section 2.3 (Figure
Figure~\ref{fig-scatterPlot}) demonstrated a generally positive
correlation between ticket sales and redemptions, suggesting that most
tickets purchased are used. However, discrepancies where sales exceed
redemptions could be linked to bulk purchases, advanced ticket sales, or
unredeemed tickets due to factors like adverse weather conditions. This
observation is consistent with the analysis by Kopystynski (1979), who
highlighted the complexities of integrating ferry systems into urban
transit networks. Understanding these discrepancies is critical for
improving ticketing policies and service offerings to better match
supply with demand.

The box plot analysis in Section 2.4 (Figure Figure~\ref{fig-boxPlot})
provided a comparative view of the variability in sales and redemption
counts. The presence of outliers, particularly during periods of high
sales but lower-than-expected redemptions, could indicate anomalies such
as large-scale events or disruptions in service. Such variability
highlights the need for a more robust system to handle unexpected
fluctuations in demand, as well as a deeper understanding of passenger
behavior. This finding resonates with the work of Roseman (2020), who
explored the unpredictable nature of ferry commutes and the challenges
faced by passengers in adapting to these changes.

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

Despite the insights gained, several limitations in the current analysis
should be addressed. First, the dataset does not capture the full
spectrum of ticket usage behavior, such as the impact of promotional
offers or the use of multi-ride passes, which could lead to an
overestimation or underestimation of actual ferry usage. Additionally,
the analysis does not incorporate external factors such as weather
conditions or special events that significantly influence ferry
operations, as noted in prior studies like McIlwraith (1984).

To address these limitations, future research should consider
incorporating supplementary data sources, such as weather information,
event schedules, and passenger feedback, to provide a more comprehensive
view of the factors affecting ferry usage. Furthermore, a comparative
study with other modes of transportation in Toronto could provide
valuable insights into the role of the ferry system within the broader
urban transit network. This would help identify areas for service
improvement and inform strategies to enhance passenger experience and
operational efficiency.

\newpage

\appendix

\section{Appendix}\label{sec-appendix}

\subsection{Dataset and Graph Sketches}\label{sec-appendix-sketches}

Sketches depicting both the desired dataset and the graphs generated in
this analysis are available in the GitHub Repository.

\subsection{Data Cleaning}\label{sec-appendix-cleaning}

The data cleaning process involved filtering out some of the columns
from the raw dataset and renaming some of the data entries for clarity
and simplicity.

\newpage

\section*{References}\label{sec-references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-kopystynski1979potential}
Kopystynski, Adrian Danie. 1979. {``The Potential of Passenger Ferries
in an Urban Transit System.''} PhD thesis, University of British
Columbia.

\bibitem[\citeproctext]{ref-mcilwraith1984transportation}
McIlwraith, Thomas F. 1984. {``TRANSPORTATION IN OLD ONTARIO.''}
\emph{American Review of Canadian Studies} 14 (2): 177--92.

\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2023. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-roseman2020precarious}
Roseman, Sharon R. 2020. {``Precarious Mobility and Spectacle Dwelling
on a Ferry Commute.''} \emph{Applied Mobilities} 5 (1): 86--102.

\bibitem[\citeproctext]{ref-varga1987toronto}
Varga, Steve. 1987. \emph{Toronto Islands}. Toronto Field Naturalists.

\bibitem[\citeproctext]{ref-wickham2016ggplot2}
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\bibitem[\citeproctext]{ref-rohan}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019a.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-wickham2019tidyverse}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romuald François, Garrett Grolemund, et al. 2019b.
{``Welcome to the Tidyverse.''} \emph{Journal of Open Source Software}.
\url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-wickham2023readr}
Wickham, Hadley, Jim Hester, and Jennifer Bryan. 2023. \emph{Readr: Read
Rectangular Text Data}. \url{https://CRAN.R-project.org/package=readr}.

\bibitem[\citeproctext]{ref-xie2023knitr}
Xie, Yihui. 2023. {``Knitr: A General-Purpose Package for Dynamic Report
Generation in r.''} \url{https://yihui.org/knitr/}.

\end{CSLReferences}

\end{document}